---
title: "World Cup Players"
output: 
  flexdashboard::flex_dashboard:
    self_contained: false
    theme: spacelab
---


Summary {data-orientation=rows}
=========================================
Row {data-height=300}
-------------------------------------

```{r, echo=FALSE}
suppressMessages(library(dplyr))
suppressMessages(library(tidyr))
suppressMessages(library(readxl))
suppressMessages(library(tinytex))
suppressMessages(library(plotly))
suppressMessages(library(lubridate))
suppressMessages(library(htmltools))
suppressMessages(library(flexdashboard))
suppressMessages(library(DT))
suppressMessages(library(scales))

WCwinners <- read_excel("footiedata/WC_Winners.xlsx", col_types = "guess")

Domestics <- WCwinners %>% filter(Domestic == "Yes")
NonDomestics <- WCwinners %>% filter(Domestic == "No")
WinDomestics <- WCwinners %>% filter(Finished == 1 & Domestic == "Yes")
LoseDomestics <- WCwinners %>% filter(Finished ==2 & Domestic == "No")
StarterDomestic <- WCwinners %>% filter(Starter == "Starter" & Domestic == "Yes")
SubDomestic <- WCwinners %>% filter(Starter == "Sub" & Domestic == "Yes")
WinStartDomestic <- WCwinners %>% filter(Finished ==1, Starter == "Starter", Domestic == "Yes")

TotalPlayers <- nrow(WCwinners)
TotalDomestics <- nrow(Domestics)
TotalWinDomestic <- nrow(WinDomestics)
TotalStartDomestic <- nrow(StarterDomestic)
TotalWinStartDomestics <- nrow(WinStartDomestic)

TotalDomesticPercent <- percent(TotalDomestics/TotalPlayers)
TotalWinDomesticPercent <- percent(TotalWinDomestic/TotalPlayers)
TotalStarterDomesticPercent <- percent(TotalStartDomestic/TotalPlayers)
TotalWinStartDomesticsPercent <- percent(TotalWinStartDomestics/TotalPlayers)

```

### WC Final Players from Home League

```{r}
gauge(TotalDomesticPercent, min = 0, max = 100, symbol = '%')
```

### WC Final Starters from Home League

```{r}
gauge(TotalStarterDomesticPercent, min = 0, max = 100, symbol = '%')
```

### WC Final Winners from Home League

```{r}
gauge(TotalWinDomesticPercent, min = 0, max = 100, symbol = '%')
```

### WC Final Winners Starting and from Home League

```{r}
gauge(TotalWinStartDomesticsPercent, min = 0, max = 100, symbol = '%')
```


Row {data-height=700}
-------------------------------------
### Domestic League Representation at WC Final

```{r, echo=FALSE}
Bardata <-
  WCwinners%>%
  group_by(Club_League, Starter)%>%
  summarise(Total = sum(Count))


LeagueBar <-
  
  plot_ly(
    data = Bardata,
    x = ~Club_League,
    y = ~Total,
    type = 'bar',
    color = ~Starter
  )%>%

  layout(barmode = 'stack')


LeagueBar
```

USA 1994
=========================================
Column {data-width=500}
-------------------------------------
WINNER BRAZIL

```{r}
TeamFilter <- function(x,y){
    n <-
      filter(WCwinners, Country == x, Year == y)
}
```

```{r}
Brazil94 <-
  TeamFilter('Brazil', 1994)
```


```{r}
Brazil94Total <- nrow(Brazil94)
Brazil94TotalDomestics <- Brazil94 %>% filter(Domestic == "Yes")
Brazil94StartDomestics <- Brazil94 %>% filter(Domestic == "Yes", Starter == "Starter")
Brazil94percentDomestics <- percent(nrow(Brazil94TotalDomestics)/nrow(Brazil94))
Brazil94percentStartDomestics <- percent(nrow(Brazil94StartDomestics)/11)
```


### Percent of Total Players from Home League

```{r}
gauge(Brazil94percentDomestics, min = 0, max = 100, symbol = '%')
```

### Percent of Starters from Home League

```{r}
gauge(Brazil94percentStartDomestics, min = 0, max = 100, symbol = '%')
```

### Brazil 1994 Players Club Leagues

```{r}
Brazil94bardata <-
  Brazil94%>%
  group_by(Club_League, Starter)%>%
  summarise(Total = sum(Count))

Brazil94bar <-
  
  plot_ly(
    data = Brazil94bardata,
    x = ~Club_League,
    y = ~Total,
    type = 'bar',
    color = ~Starter
  )%>%

  layout(barmode = 'stack')


Brazil94bar
```

Column {data-width=500}
-------------------------------------

LOSER ITALY

```{r}
Italy94 <-
  TeamFilter('Italy', 1994)
```

```{r}
Italy94Total <- nrow(Italy94)
Italy94TotalDomestics <- Italy94 %>% filter(Domestic == "Yes")
Italy94StartDomestics <- Italy94 %>% filter(Domestic == "Yes", Starter == "Starter")
Italy94percentDomestics <- percent(nrow(Italy94TotalDomestics)/nrow(Italy94))
Italy94percentStartDomestics <- percent(nrow(Italy94StartDomestics)/11)
```

### Percent of Total Players from Home League

```{r}
gauge(Italy94percentDomestics, min = 0, max = 100, symbol = '%')
```

### Percent of Starters from Home League

```{r}
gauge(Italy94percentStartDomestics, min = 0, max = 100, symbol = '%')
```

### Italy 1994 Players Club Leagues

```{r}
Italy94bardata <-
  Italy94%>%
  group_by(Club_League, Starter)%>%
  summarise(Total = sum(Count))

Italy94bar <-
  
  plot_ly(
    data = Italy94bardata,
    x = ~Club_League,
    y = ~Total,
    type = 'bar',
    color = ~Starter
  )%>%

  layout(barmode = 'stack')


Italy94bar
```

FRANCE 1998
=========================================
Column {data-width=500}
-------------------------------------
WINNER FRANCE

```{r}
TeamFilter <- function(x,y){
    n <-
      filter(WCwinners, Country == x, Year == y)
}
```

```{r}
France98 <-
  TeamFilter('France', 1998)
```


```{r}
France98Total <- nrow(Brazil94)
France98TotalDomestics <- France98 %>% filter(Domestic == "Yes")
France98StartDomestics <- France98 %>% filter(Domestic == "Yes", Starter == "Starter")
France98percentDomestics <- percent(nrow(France98TotalDomestics)/nrow(France98))
France98percentStartDomestics <- percent(nrow(France98StartDomestics)/11)
```


### Percent of Total Players from Home League

```{r}
gauge(France98percentDomestics, min = 0, max = 100, symbol = '%')
```

### Percent of Starters from Home League

```{r}
gauge(France98percentStartDomestics, min = 0, max = 100, symbol = '%')
```

### France 1998 Players Club Leagues

```{r}
France98bardata <-
  France98%>%
  group_by(Club_League, Starter)%>%
  summarise(Total = sum(Count))

France98bar <-
  
  plot_ly(
    data = France98bardata,
    x = ~Club_League,
    y = ~Total,
    type = 'bar',
    color = ~Starter
  )%>%

  layout(barmode = 'stack')


France98bar
```

Column {data-width=500}
-------------------------------------

LOSER BRAZIL

```{r}
Brazil98 <-
  TeamFilter('Brazil', 1998)
```

```{r}
Brazil98Total <- nrow(Brazil98)
Brazil98TotalDomestics <- Brazil98 %>% filter(Domestic == "Yes")
Brazil98StartDomestics <- Brazil98 %>% filter(Domestic == "Yes", Starter == "Starter")
Brazil98percentDomestics <- percent(nrow(Brazil98TotalDomestics)/nrow(Brazil98))
Brazil98percentStartDomestics <- percent(nrow(Brazil98StartDomestics)/11)
```

### Percent of Total Players from Home League

```{r}
gauge(Brazil98percentDomestics, min = 0, max = 100, symbol = '%')
```

### Percent of Starters from Home League

```{r}
gauge(Brazil98percentStartDomestics, min = 0, max = 100, symbol = '%')
```

### Brazil 1998 Players Club Leagues

```{r}
Brazil98bardata <-
  Brazil98%>%
  group_by(Club_League, Starter)%>%
  summarise(Total = sum(Count))

Brazil98bar <-
  
  plot_ly(
    data = Brazil98bardata,
    x = ~Club_League,
    y = ~Total,
    type = 'bar',
    color = ~Starter
  )%>%

  layout(barmode = 'stack')


Brazil98bar
```

JAPAN & KOREA 2002
=========================================
Column {data-width=500}
-------------------------------------
WINNER BRAZIL

```{r}
TeamFilter <- function(x,y){
    n <-
      filter(WCwinners, Country == x, Year == y)
}
```

```{r}
Brazil02 <-
  TeamFilter('Brazil', 2002)
```


```{r}
Brazil02Total <- nrow(Brazil02)
Brazil02TotalDomestics <- Brazil02 %>% filter(Domestic == "Yes")
Brazil02StartDomestics <- Brazil02 %>% filter(Domestic == "Yes", Starter == "Starter")
Brazil02percentDomestics <- percent(nrow(Brazil02TotalDomestics)/nrow(Brazil02))
Brazil02percentStartDomestics <- percent(nrow(Brazil02StartDomestics)/11)
```


### Percent of Total Players from Home League

```{r}
gauge(Brazil02percentDomestics, min = 0, max = 100, symbol = '%')
```

### Percent of Starters from Home League

```{r}
gauge(Brazil02percentStartDomestics, min = 0, max = 100, symbol = '%')
```

### Brazil 2002 Players Club Leagues

```{r}
Brazil02bardata <-
  Brazil02%>%
  group_by(Club_League, Starter)%>%
  summarise(Total = sum(Count))

Brazil02bar <-
  
  plot_ly(
    data = Brazil02bardata,
    x = ~Club_League,
    y = ~Total,
    type = 'bar',
    color = ~Starter
  )%>%

  layout(barmode = 'stack')


Brazil02bar
```

Column {data-width=500}
-------------------------------------

LOSER GERMANY

```{r}
Germany02 <-
  TeamFilter('Germany', 2002)
```

```{r}
Germany02Total <- nrow(Germany02)
Germany02TotalDomestics <- Germany02 %>% filter(Domestic == "Yes")
Germany02StartDomestics <- Germany02 %>% filter(Domestic == "Yes", Starter == "Starter")
Germany02percentDomestics <- percent(nrow(Germany02TotalDomestics)/nrow(Germany02))
Germany02percentStartDomestics <- percent(nrow(Germany02StartDomestics)/11)
```

### Percent of Total Players from Home League

```{r}
gauge(Germany02percentDomestics, min = 0, max = 100, symbol = '%')
```

### Percent of Starters from Home League

```{r}
gauge(Germany02percentStartDomestics, min = 0, max = 100, symbol = '%')
```

### Germany 2002 Players Club Leagues

```{r}
Germany02bardata <-
  Germany02%>%
  group_by(Club_League, Starter)%>%
  summarise(Total = sum(Count))

Germany02bar <-
  
  plot_ly(
    data = Germany02bardata,
    x = ~Club_League,
    y = ~Total,
    type = 'bar',
    color = ~Starter
  )%>%

  layout(barmode = 'stack')


Germany02bar
```

GERMANY 2006
=========================================
Column {data-width=500}
-------------------------------------
WINNER ITALY

```{r}
Italy06 <-
  TeamFilter('Italy', 2006)
```


```{r}
Italy06Total <- nrow(Italy06)
Italy06TotalDomestics <- Italy06 %>% filter(Domestic == "Yes")
Italy06StartDomestics <- Italy06 %>% filter(Domestic == "Yes", Starter == "Starter")
Italy06percentDomestics <- percent(nrow(Italy06TotalDomestics)/nrow(Italy06))
Italy06percentStartDomestics <- percent(nrow(Italy06StartDomestics)/11)
```


### Percent of Total Players from Home League

```{r}
gauge(Italy06percentDomestics, min = 0, max = 100, symbol = '%')
```

### Percent of Starters from Home League

```{r}
gauge(Italy06percentStartDomestics, min = 0, max = 100, symbol = '%')
```

### Italy 2006 Players Club Leagues

```{r}
Italy06bardata <-
  Italy06%>%
  group_by(Club_League, Starter)%>%
  summarise(Total = sum(Count))

Italy06bar <-
  
  plot_ly(
    data = Italy06bardata,
    x = ~Club_League,
    y = ~Total,
    type = 'bar',
    color = ~Starter
  )%>%

  layout(barmode = 'stack')


Italy06bar
```

Column {data-width=500}
-------------------------------------

LOSER FRANCE

```{r}
France06 <-
  TeamFilter('France', 2006)
```

```{r}
France06Total <- nrow(France06)
France06TotalDomestics <- France06 %>% filter(Domestic == "Yes")
France06StartDomestics <- France06 %>% filter(Domestic == "Yes", Starter == "Starter")
France06percentDomestics <- percent(nrow(France06TotalDomestics)/nrow(France06))
France06percentStartDomestics <- percent(nrow(France06StartDomestics)/11)
```

### Percent of Total Players from Home League

```{r}
gauge(France06percentDomestics, min = 0, max = 100, symbol = '%')
```

### Percent of Starters from Home League

```{r}
gauge(France06percentStartDomestics, min = 0, max = 100, symbol = '%')
```

### France 2006 Players Club Leagues

```{r}
France06bardata <-
  France06%>%
  group_by(Club_League, Starter)%>%
  summarise(Total = sum(Count))

France06bar <-
  
  plot_ly(
    data = France06bardata,
    x = ~Club_League,
    y = ~Total,
    type = 'bar',
    color = ~Starter
  )%>%

  layout(barmode = 'stack')


France06bar
```

SOUTH AFRICA 2010
=========================================
Column {data-width=500}
-------------------------------------
WINNER SPAIN

```{r}
Spain10 <-
  TeamFilter('Spain', 2010)
```


```{r}
Spain10Total <- nrow(Spain10)
Spain10TotalDomestics <- Spain10 %>% filter(Domestic == "Yes")
Spain10StartDomestics <- Spain10 %>% filter(Domestic == "Yes", Starter == "Starter")
Spain10percentDomestics <- percent(nrow(Spain10TotalDomestics)/nrow(Spain10))
Spain10percentStartDomestics <- percent(nrow(Spain10StartDomestics)/11)
```


### Percent of Total Players from Home League

```{r}
gauge(Spain10percentDomestics, min = 0, max = 100, symbol = '%')
```

### Percent of Starters from Home League

```{r}
gauge(Spain10percentStartDomestics, min = 0, max = 100, symbol = '%')
```

### Spain 2010 Players Club Leagues

```{r}
Spain10bardata <-
  Spain10%>%
  group_by(Club_League, Starter)%>%
  summarise(Total = sum(Count))

Spain10bar <-
  
  plot_ly(
    data = Spain10bardata,
    x = ~Club_League,
    y = ~Total,
    type = 'bar',
    color = ~Starter
  )%>%

  layout(barmode = 'stack')


Spain10bar
```

Column {data-width=500}
-------------------------------------

LOSER NETHERLANDS

```{r}
Netherlands10 <-
  TeamFilter('Netherlands', 2010)
```

```{r}
Netherlands10Total <- nrow(Netherlands10)
Netherlands10TotalDomestics <- Netherlands10 %>% filter(Domestic == "Yes")
Netherlands10StartDomestics <- Netherlands10 %>% filter(Domestic == "Yes", Starter == "Starter")
Netherlands10percentDomestics <- percent(nrow(Netherlands10TotalDomestics)/nrow(Netherlands10))
Netherlands10percentStartDomestics <- percent(nrow(Netherlands10StartDomestics)/11)
```

### Percent of Total Players from Home League

```{r}
gauge(Netherlands10percentDomestics, min = 0, max = 100, symbol = '%')
```

### Percent of Starters from Home League

```{r}
gauge(Netherlands10percentStartDomestics, min = 0, max = 100, symbol = '%')
```

### Netherlands 2010 Players Club Leagues

```{r}
Netherlands10bardata <-
  Netherlands10%>%
  group_by(Club_League, Starter)%>%
  summarise(Total = sum(Count))

Netherlands10bar <-
  
  plot_ly(
    data = Netherlands10bardata,
    x = ~Club_League,
    y = ~Total,
    type = 'bar',
    color = ~Starter
  )%>%

  layout(barmode = 'stack')


Netherlands10bar
```

BRAZIL 2014
=========================================
Column {data-width=500}
-------------------------------------
WINNER GERMANY

```{r}
Germany14 <-
  TeamFilter('Germany', 2014)
```


```{r}
Germany14Total <- nrow(Germany14)
Germany14TotalDomestics <- Germany14 %>% filter(Domestic == "Yes")
Germany14StartDomestics <- Germany14 %>% filter(Domestic == "Yes", Starter == "Starter")
Germany14percentDomestics <- percent(nrow(Germany14TotalDomestics)/nrow(Germany14))
Germany14percentStartDomestics <- percent(nrow(Germany14StartDomestics)/11)
```


### Percent of Total Players from Home League

```{r}
gauge(Germany14percentDomestics, min = 0, max = 100, symbol = '%')
```

### Percent of Starters from Home League

```{r}
gauge(Germany14percentStartDomestics, min = 0, max = 100, symbol = '%')
```

### Germany 2014 Players Club Leagues

```{r}
Germany14bardata <-
  Germany14%>%
  group_by(Club_League, Starter)%>%
  summarise(Total = sum(Count))

Germany14bar <-
  
  plot_ly(
    data = Germany14bardata,
    x = ~Club_League,
    y = ~Total,
    type = 'bar',
    color = ~Starter
  )%>%

  layout(barmode = 'stack')


Germany14bar
```

Column {data-width=500}
-------------------------------------

LOSER ARGENTINA

```{r}
Argentina14 <-
  TeamFilter('Argentina', 2014)
```

```{r}
Argentina14Total <- nrow(Argentina14)
Argentina14TotalDomestics <- Argentina14 %>% filter(Domestic == "Yes")
Argentina14StartDomestics <- Argentina14 %>% filter(Domestic == "Yes", Starter == "Starter")
Argentina14percentDomestics <- percent(nrow(Argentina14TotalDomestics)/nrow(Argentina14))
Argentina14percentStartDomestics <- percent(nrow(Argentina14StartDomestics)/11)
```

### Percent of Total Players from Home League

```{r}
gauge(Argentina14percentDomestics, min = 0, max = 100, symbol = '%')
```

### Percent of Starters from Home League

```{r}
gauge(Argentina14percentStartDomestics, min = 0, max = 100, symbol = '%')
```

### Argentina 2014 Players Club Leagues

```{r}
Argentina14bardata <-
  Argentina14%>%
  group_by(Club_League, Starter)%>%
  summarise(Total = sum(Count))

Argentina14bar <-
  
  plot_ly(
    data = Argentina14bardata,
    x = ~Club_League,
    y = ~Total,
    type = 'bar',
    color = ~Starter
  )%>%

  layout(barmode = 'stack')


Argentina14bar
```

RUSSIA 2018
=========================================
Column {data-width=500}
-------------------------------------
WINNER FRANCE

```{r}
France18 <-
  TeamFilter('France', 2018)
```


```{r}
France18Total <- nrow(France18)
France18TotalDomestics <- France18 %>% filter(Domestic == "Yes")
France18StartDomestics <- France18 %>% filter(Domestic == "Yes", Starter == "Starter")
France18percentDomestics <- percent(nrow(France18TotalDomestics)/nrow(France18))
France18percentStartDomestics <- percent(nrow(France18StartDomestics)/11)
```


### Percent of Total Players from Home League

```{r}
gauge(France18percentDomestics, min = 0, max = 100, symbol = '%')
```

### Percent of Starters from Home League

```{r}
gauge(France18percentStartDomestics, min = 0, max = 100, symbol = '%')
```

### France 2018 Players Club Leagues

```{r}
France18bardata <-
  France18%>%
  group_by(Club_League, Starter)%>%
  summarise(Total = sum(Count))

France18bar <-
  
  plot_ly(
    data = France18bardata,
    x = ~Club_League,
    y = ~Total,
    type = 'bar',
    color = ~Starter
  )%>%

  layout(barmode = 'stack')


France18bar
```

Column {data-width=500}
-------------------------------------

LOSER CROATIA

```{r}
Croatia18 <-
  TeamFilter('Croatia', 2018)
```

```{r}
Croatia18Total <- nrow(Croatia18)
Croatia18TotalDomestics <- Croatia18 %>% filter(Domestic == "Yes")
Croatia18StartDomestics <- Croatia18 %>% filter(Domestic == "Yes", Starter == "Starter")
Croatia18percentDomestics <- percent(nrow(Croatia18TotalDomestics)/nrow(Croatia18))
Croatia18percentStartDomestics <- percent(nrow(Croatia18StartDomestics)/11)
```

### Percent of Total Players from Home League

```{r}
gauge(Croatia18percentDomestics, min = 0, max = 100, symbol = '%')
```

### Percent of Starters from Home League

```{r}
gauge(Croatia18percentStartDomestics, min = 0, max = 100, symbol = '%')
```

### Croatia 2018 Players Club Leagues

```{r}
Croatia18bardata <-
  Argentina14%>%
  group_by(Club_League, Starter)%>%
  summarise(Total = sum(Count))

Croatia18bar <-
  
  plot_ly(
    data = Croatia18bardata,
    x = ~Club_League,
    y = ~Total,
    type = 'bar',
    color = ~Starter
  )%>%

  layout(barmode = 'stack')


Croatia18bar
```

DATA
=========================================

```{r}
datatable(WCwinners)
```

